---
title: "The Toronto Analytics Job Landscape"
author: "Summerhill"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  md_document:
    variant: markdown_github
  html_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=10, fig.path='Figs/',
                      warning=FALSE,message=FALSE,error=FALSE,include=TRUE,echo=TRUE)

set.seed(42)

# a note on options: 
# include=TRUE means that output of the code is shown in the document (intended for graphics).
# echo=TRUE means the code is shown in the document (intended for code that might be interesting for a reader).
# message and warning are for the text for loading libraries or if a function fails
# echo=TRUE means the code is not shown in the final document
# http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html
# http://rmarkdown.rstudio.com/developer_parameterized_reports.html
# https://guides.github.com/features/mastering-markdown/

```
- This document was rendered last on `r Sys.Date()`

##*THIS PROJECT IS STILL UNDER CONSTRUCTION*
The intention will be to mask the code as the project approaches completion.

##Executive Summary
- This project was produced for the Text Analytics Workshop for the Winter 2018 Masters of Management Analytics Cohort at Queen's University
- The goal from the outset was to use text analytics techniques developed in class to examine jobs companies have posted on Indeed in Toronto
and employ techniques discussed in class including document clustering, topic modelling, and visualization.

##Project Rationale
- A open sourced project working with real world data was desired
- Other projects can be found scraping DS/Analytics jobs from Indeed. Typically word frequencies for keywords like Python or Hadoop are calculated
- Moving beyond that, we were interested in topic modelling and how the choice of words signals relationships between roles
- Job postings fit the 'bag of words' or ngram approach taught in class. Not many employers say **"We don't want someone who knows Python"**

```{r Import Libraries}
library(feather)
library(tidyverse)
library(tidytext)
library(tm)
library(wordcloud)
library(widyr)
library(ggraph)
library(igraph)
library(knitr)
library(ggridges)
library(purrrlyr)
```

##Gathering Data
- Beautiful Soup & Selenium were used in Python to access [Indeed](https://www.indeed.ca/jobs?q=analytics&l=Toronto&start=10 "Indeed:Analytics Jobs in Toronto") and scrape unsponsored job titles, companies, and postings
- `later number` unique jobs were scraped from the search terms: `analytics`,`etc`....
- Jobs were passed from Python to R using [Feather](https://blog.rstudio.com/2016/03/29/feather/ "Feather: A Fast On-Disk Format for Data Frames for R and Python, powered by Apache Arrow")

```{r Import Data,echo=TRUE}
rm(list=ls())
#list our data files
searches <- c("analytics",
                 "data analyst",
                 "data scientist",
                 "analytics strategy",
                 "data insights",
                 "marketing analytics",
                 "analytics reporting",
                 "machine learning",
                 "business intelligence")

files <- paste("data/feather/",searches,".feather",sep="")

#read and collapse to data frame
datalist <- lapply(as.list(files),function(x){read_feather(x)})
test <- datalist[[1]]
data <- bind_rows(datalist,.id="search")
rm(datalist)

#fix quotations in column names
names(data) <- c("search","company","text","titles","urls")
data <- data %>% select(company,titles,text,search,urls)

#check if we have redundant jobs
sum(duplicated(data[,2:4]))

#examine the uniqueness of our data
NumJobs <- n_distinct(data$urls)

#reduce to distinct jobs and clean up search column
data <- data[!duplicated(data$urls),]
data$search <- plyr::mapvalues(data$search,
                               from=unique(data$search),
                               to=searches)
```
- Our data returned `NumJobs` unique jobs within our search.
- It's clear a considerable amount of cleaning is in order 

```{r preprocessing}
RemovePattern <- function(vector,pattern){gsub(pattern=pattern,replacement=" ",vector)}
data <- dmap(data,RemovePattern,"\n")
data <- dmap(data,RemovePattern,"\\(")   
data <- dmap(data,RemovePattern,"\\)")
data <- dmap(data,RemovePattern,"\\{[a-zA-Z0-9]\\}")
```

```{r Jobs Found}
#investigate redundant jobs. Should return 200/each if they are all unique.

rollup <- data %>%
     group_by(search) %>%
     summarize(NumberUniquePostings=n())

str(rollup)

#sort by search order
kable(left_join(data.frame(search=searches),rollup,by="search"))
```

- We expect 200 jobs for each result, and removing the duplicate jobs in the order they were searched.
- Interestingly, searching 200 jobs in analytics returns only 113 unique jobs, some redundancy exists.
- As we search overlapping terms, data sciencist, data insights, fewer and fewer unique jobs are returned
- Interestingly, each additional search term returns a surprising amount of new jobs. 75 jobs are shown for machine learning that were not found for data scientist, a fairly similar field.
- Business Intelligence seems to be fairly lateral to other search terms, returning many unique jobs

```{r}
kable(head(data %>% group_by(search,titles) %>% count(sort=TRUE),10))
```
```{r}
kable(head(data %>% count(company,sort=TRUE),10))
```

```{r Remove empty jobs}
#how long are our jobs
data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(urls) %>%
     count(search,urls,sort=TRUE) %>%
     filter(n<3000) %>%
     ggplot(aes(x=n,y=search))+geom_density_ridges()
```

- We see that there are alot of 0 information jobs, postings with only a few words, let's look at what some of those are.

```{r remove outliers}
MinWords <- 300

empty_urls <- data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(urls) %>%
     count(urls,sort=TRUE) %>%
     filter(n < MinWords)

# data %>% 
#      filter(urls %in% empty_urls$urls) %>%
#      head()

data <- data %>% 
     filter(!urls %in% empty_urls$urls)
     
```

- We see here a variety of failed scrapings, or 0 info postings


```{r long postings}
long_postings <- data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(urls) %>%
     count(urls,sort=TRUE) %>%
     filter(n >3000)

```

```{r}
unigrams <- data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(unigrams) %>%
     count(unigrams,sort=TRUE)

wordcloud(unigrams$unigrams,unigrams$n,max.words=50)
```


- Looking at a simple word frequency, we see out of the box our data is very messy
- The boiler plate at the end of each job posting, encouraging people to apply, discussing company acolades and culture distort our analysis. Let's spend some time cleaning up *common stopwords*, *job specific words* and *html*

```{r Process unigrams Data}
#what words to avoid
stop <- read.csv("stopwords.csv",stringsAsFactors = FALSE)
stop <- rbind(stop,data.frame(words=stopwords("en")))


#process n-grams
unigrams <- data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(unigrams) %>%
     filter(!unigrams %in% stop$words) %>%
     count(unigrams,sort=TRUE)

#visualize
wordcloud(unigrams$unigrams,unigrams$n,max.words=50)
```

- We are starting to look better. Let's take a look at our bigrams.
```{r Process bigrams}
#look a bi-grams
bigrams_totals <- data %>%
     unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
     separate(col=tokens,into=c("word1","word2"),sep=" ") %>%
     filter(!word1 %in% stop$words, !word2 %in% stop$words) %>%
     unite(tokens,word1,word2,sep=" ") %>%
     count(tokens,sort=TRUE)

#kable(head(bigrams_totals,20))

wordcloud(bigrams_totals$tokens,bigrams_totals$n,max.words=10)
```



```{r Unigram frequency}

#determine how frequent each word occurs across job postings. Don't skip stop words yet.
unigrams_freq <- data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     select(urls,tokens) %>%
     distinct() %>%
     filter(!tokens %in% stop$words) %>%
     group_by(tokens) %>%
     count(tokens,sort=TRUE) %>%
     ungroup() %>%
     mutate(frequency=n/NumJobs)

kable(head(unigrams_freq,20))
```


```{r}

#determine how frequent each bigram is across job postings. Don't skip stop words yet.
bigrams_freq <- data %>%
     unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
     select(urls,tokens) %>%
     distinct() %>%
     separate(col=tokens,into=c("word1","word2"),sep=" ") %>%
     filter(!word1 %in% stop$words, !word2 %in% stop$words) %>%
     unite(tokens,word1,word2,sep=" ") %>%
     group_by(tokens) %>%
     count(tokens,sort=TRUE) %>%
     ungroup() %>%
     mutate(frequency=n/NumJobs)

kable(head(bigrams_freq,20))

```



```{r pairwise count}

skills <- c("sas","python","sql","r","spss",
            "hadoop","spark","java","scala","aws","c++",
            "excel","powerpoint","mapreduce",
          "qlik","tableau","microstrategy","azure")

data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     filter(tokens %in% skills) %>%
     pairwise_count(tokens,urls,sort=TRUE) %>%
     head()

data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     filter(tokens %in% skills) %>%
     ggplot(aes(x=tokens)) +
     geom_bar()
```





```{r pairwise correlation}
#pairwise correlation
data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     filter(tokens %in% skills) %>%
     pairwise_cor(tokens,urls,sort=TRUE) %>%
     filter(correlation > .2) %>%
     graph_from_data_frame() %>%
     ggraph(layout = "fr") +
     geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
     geom_node_point(color = "lightblue", size = 5) +
     geom_node_text(aes(label = name), repel = TRUE) +
     theme_void()

```

```{r clustering}

#Frequency filters
minFreq = 0.01
maxFreq = 0.80

#filtering out stopwords and infrequent/frequent words from unigrams
unigram_clean <- data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     select(urls,tokens) %>%
     distinct() %>%
     filter(!tokens %in% stop$words) %>%
     group_by(tokens) %>%
     count(tokens,sort=TRUE) %>%
     mutate(frequency=n/NumJobs) %>%
     filter(frequency> minFreq & frequency< maxFreq)

#creating identifier
data$ID <- paste(data$company,data$titles,sep="_")

#creating clean unigrams DTM
unigram_cluster <- data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     select(ID,tokens) %>%
     filter(tokens %in% unigram_clean$tokens) %>% 
     count(ID,tokens,sort=TRUE) %>%
     ungroup() %>%
     spread(tokens,n,fill='0') 

#filtering out stopwords and infrequent/frequent words from bigrams
bigram_clean <- data %>%
     unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
     select(urls,tokens) %>%
     distinct() %>%
     separate(col=tokens,into=c("word1","word2"),sep=" ") %>%
     filter(!word1 %in% stop$words, !word2 %in% stop$words) %>%
     unite(tokens,word1,word2,sep=" ") %>%
     group_by(tokens) %>%
     count(tokens,sort=TRUE) %>%
     ungroup() %>%
     mutate(frequency=n/NumJobs) %>%
     filter(frequency> minFreq & frequency< maxFreq)

#creating clean bigrams DTM
bigram_cluster <- data %>%
     unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
     select(ID,tokens) %>%
     filter(tokens %in% bigram_clean$tokens) %>% 
     count(ID,tokens,sort=TRUE) %>%
     ungroup() %>%
     spread(tokens,n,fill='0') 

#Joining unigrams and bigrams cluster data together 

data_cluster <- left_join(unigram_cluster,bigram_cluster)

#transposing so terms become the rows
data_cluster <- t(data_cluster)
data_cluster <- as.data.frame(data_cluster)
#reducing sparsity of TDM before clustering
dim(data_cluster)
data_clean <- removeSparseTerms(data_cluster,sparse = 0.80)
data_m <- as.matrix(data_clean)
data_df <- as.data.frame(data_m)

#looking at most popular words, min freq of 20
findFreqTerms(data_cluster,lowfreq=20)

#accumulator of clustering results
clust_results <- data.frame()

#run kmeans for all clusters up to 15
set.seed(100)
for(i in 1:20) {
     k_clust <- kmeans(data_cluster[,-1], centers=i, iter.max =100)
     #Combine cluster number and cluster together, write to df
     clust_results <- rbind(clust_results,cbind(i,k_clust$tot.withinss))
}
names(clust_results) <- c("cluster","results")

#scree elbow plot
ggplot(clust_results,aes(x=cluster,y=results,group=1))+
geom_point()+geom_line()

#if you want specific points look at scaling the axes then use seq(1:10)

## above plot shows that at about 3 clusters, any additional clusters would lose significance

#need to create principal components, regular kmeans will kick out an error because it uses princomp function
pc <- prcomp(data_cluster)
plot(pc,xlab="Principal Components")
summary(pc)
comp <- data.frame(pc$x[,1:12])
plot(comp)

#Let's look at 3 clusters
kmeans3 <- kmeans(comp,centers=3)

#Plot clusters
library(cluster)
clusplot(comp, kmeans3$cluster, color=TRUE, shade=TRUE,
         labels=2, lines=0, main="K-means cluster plot")

###hierarchical clustering

#calculate distance
data_dist <- dist(data_df,method="euclidean")

#creating and plotting hierarchical cluster
hc <- hclust(data_dist)
plot(hc)
rect.hclust(hc,k=8,border="red")
```



```{r Clustering with dplyr}

# #filtering out stopwords and infrequent/frequent words from unigrams
# unigram_clean <- data %>%
#      unnest_tokens(token="words",output="tokens",input=text) %>%
#      select(urls,tokens) %>%
#      distinct() %>%
#      filter(!tokens %in% stop$words) %>%
#      group_by(tokens) %>%
#      count(tokens,sort=TRUE) %>%
#      mutate(frequency=n/NumJobs) %>%
#      filter(frequency> minFreq & frequency< maxFreq)
# 
# #creating identifier
# data$ID <- paste(data$company,data$titles,sep="_")
# 
# #creating clean unigrams TDM
# unigram_cluster <- data %>%
#      unnest_tokens(token="words",output="tokens",input=text) %>%
#      select(ID,tokens) %>%
#      filter(tokens %in% unigram_clean$tokens) %>% 
#      count(ID,tokens,sort=TRUE) %>%
#      ungroup() %>%
#      spread(tokens,n,fill='0') 
# 
# #filtering out stopwords and infrequent/frequent words from bigrams
# bigram_clean <- data %>%
#      unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
#      select(urls,tokens) %>%
#      distinct() %>%
#      separate(col=tokens,into=c("word1","word2"),sep=" ") %>%
#      filter(!word1 %in% stop$words, !word2 %in% stop$words) %>%
#      unite(tokens,word1,word2,sep=" ") %>%
#      group_by(tokens) %>%
#      count(tokens,sort=TRUE) %>%
#      ungroup() %>%
#      mutate(frequency=n/NumJobs) %>%
#      filter(frequency> minFreq & frequency< maxFreq)
# 
# #creating clean bigrams TDM
# bigram_cluster <- data %>%
#      unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
#      select(ID,tokens) %>%
#      filter(tokens %in% bigram_clean$tokens) %>% 
#      count(ID,tokens,sort=TRUE) %>%
#      ungroup() %>%
#      spread(tokens,n,fill='0') 
# 
# #Joining unigrams and bigrams cluster data together 
# 
# data_cluster <- full_join(unigram_cluster,bigram_cluster)
# 
# #do the same clustering
# 
# #when you now do this route, all your stop words should be gone.

```







