---
title: "The Toronto Analytics Job Landscape"
author: "Summerhill"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
  html_document: default
  pdf_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=5, fig.path='Figs/',
                      warning=FALSE,message=FALSE,error=FALSE,include=TRUE,echo=TRUE)

# a note on options: 
# include=TRUE means that output of the code is shown in the document (intended for graphics).
# echo=TRUE means the code is shown in the document (intended for code that might be interesting for a reader).
# message and warning are for the text for loading libraries or if a function fails
# echo=TRUE means the code is not shown in the final document
# http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html
# http://rmarkdown.rstudio.com/developer_parameterized_reports.html
# https://guides.github.com/features/mastering-markdown/

```
- This document was rendered last on `r Sys.Date()`

##*THIS PROJECT IS STILL UNDER CONSTRUCTION*
The intention will be to mask the code as the project approaches completion.

##Executive Summary
- This project was produced for the Text Analytics Workshop for the Winter 2018 Masters of Management Analytics Cohort at Queen's University
- The goal from the outset was to use text analytics techniques developed in class to examine jobs companies have posted on Indeed in Toronto
and employ techniques discussed in class including document clustering, topic modelling, and visualization.

##Project Rationale
- A open sourced project working with real world data was desired
- Other projects can be found scraping DS/Analytics jobs from Indeed. Typically word frequencies for keywords like Python or Hadoop are calculated
- Moving beyond that, we were interested in topic modelling and how the choice of words signals relationships between roles
- Job postings fit the 'bag of words' or ngram approach taught in class. Not many employers say **"We don't want someone who knows Python"**

```{r Import Libraries}
library(feather)
library(tidyverse)
library(tidytext)
library(tm)
library(wordcloud)
library(widyr)
library(ggraph)
library(igraph)
```

##Gathering Data
- Beautiful Soup & Selenium were used in Python to access [Indeed](https://www.indeed.ca/jobs?q=analytics&l=Toronto&start=10 "Indeed:Analytics Jobs in Toronto") and scrape unsponsored job titles, companies, and postings
- `later number` unique jobs were scraped from the search terms: `analytics`,`etc`....
- Jobs were passed from Python to R using [Feather](https://blog.rstudio.com/2016/03/29/feather/ "Feather: A Fast On-Disk Format for Data Frames for R and Python, powered by Apache Arrow")

```{r Import Data,echo=TRUE}
rm(list=ls())
#list our data files
searches <- c("analytics",
                 "data analyst",
                 "data scientist",
                 "analytics strategy",
                 "data insights",
                 "marketing analytics",
                 "analytics reporting",
                 "machine learning",
                 "business intelligence")

files <- paste("data/feather/",searches,".feather",sep="")

#read and collapse to data frame
datalist <- lapply(as.list(files),function(x){read_feather(x)})
data <- bind_rows(datalist,.id="search")
rm(datalist)

#fix quotations in column names
names(data) <- c("search","company","text","titles","urls")

#check if we have redundant jobs
sum(duplicated(data[,2:4]))

#reduce to distinct jobs and clean up search column
data <- data[!duplicated(data[,2:4]),]
data$search <- plyr::mapvalues(data$search,
                               from=unique(data$search),
                               to=searches)

head(data)
unique(data$search)
```

```{r Jobs Found}
#investigate redundant jobs. Should return 20/each if they are all unique.
data %>%
     group_by(search) %>%
     summarize(found_jobs=n())

```

We expect 200 jobs for each result, and removing the duplicate jobs we can begin to see some overlap. Analytics had more jobs since it was searched first, topics discussed later like BI still has alot of jobs, which suggests it may be a more distinct field.

```{r Process unigrams Data}

#what words to avoid
stops <- stopwords("en")

unigrams <- data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     group_by(tokens) %>%
     filter(!tokens %in% stops) %>%
     count(tokens,sort=TRUE)

unigrams
```






```{r Process bigrams}
#look a bi-grams
data %>%
     unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
     separate(col=tokens,into=c("word1","word2"),sep=" ") %>%
     filter(!word1 %in% stops, !word2 %in% stops) %>%
     unite(tokens,word1,word2,sep=" ") %>%
     count(tokens,sort=TRUE)
```



```{r Clean Data}

#Converting text column into VectorSource
data_source <- VectorSource(data$text)

#Creating a corpus out of the data
data_corpus <- VCorpus(data_source)

#clean data
#data_corpus <- clean_corpus(data_corpus)

```

```{r Build 1gram 2 gram Lexicons}

```

```{r Word Frequencies}

```

```{r Convert to TDM}

#Converting corpus into TDM
#data_tdm <- TermDocumentMatrix(data_corpus)
```

```{r Word Cloud}

# #convert TDM into matrix
# data_m <- as.matrix(data_tdm)
# 
# #count term frequencies and sort in descending order
# term_frequency <- sort(rowSums(data_m),decreasing = TRUE)
# term_frequency[1:15]
# 
# #create data frame of different word frequencies
# words <- data.frame(term = names(term_frequency),num=term_frequency)
# 
# #create word cloud
# wordcloud(words$term,words$num,max.words = 75, colors = "blue")

wordcloud(unigrams$tokens,unigrams$n,max.words=100)
```

```{r pairwise count}

data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     filter(tokens %in% c("sas","python","excel","powerpoint","sql",
                          "r","hadoop","spark","java","scala","aws","tableau","microstrategy")) %>%
     pairwise_count(tokens,urls,sort=TRUE)
```





```{r pairwise correlation}
#pairwise correlation
data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     filter(tokens %in% c("sas","python","excel","powerpoint","sql",
                          "r","hadoop","spark","java","scala","aws","tableau","microstrategy")) %>%
     pairwise_cor(tokens,urls,sort=TRUE)

```






