---
title: "The Toronto Analytics Job Landscape"
author: "Summerhill"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
  html_document: default
  pdf_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=5, fig.path='Figs/',
                      warning=FALSE, message=FALSE,error=FALSE,include=TRUE,echo=TRUE)

rm(list=ls())
# a note on options: 
# include=TRUE means that output of the code is shown in the document (intended for graphics).
# echo=TRUE means the code is shown in the document (intended for code that might be interesting for a reader).
# message and warning are for the text for loading libraries or if a function fails
# echo=TRUE means the code is not shown in the final document
# http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html
# http://rmarkdown.rstudio.com/developer_parameterized_reports.html
# https://guides.github.com/features/mastering-markdown/

```
- This document was rendered last on `r Sys.Date()`

##*THIS PROJECT IS STILL UNDER CONSTRUCTION*
The intention will be to mask the code as the project approaches completion.

##Executive Summary
- This project was produced for the Text Analytics Workshop for the Winter 2018 Masters of Management Analytics Cohort at Queen's University
- The goal from the outset was to use text analytics techniques developed in class to examine jobs companies have posted on Indeed in Toronto
and employ techniques discussed in class including document clustering, topic modelling, and visualization.


```{r Import Libraries}
library(feather)
library(tidyverse)
library(tidytext)
library(tm)
```

##Gathering Data
- Beautiful Soup & Selenium were used in Python to access [Indeed](https://www.indeed.ca/jobs?q=analytics&l=Toronto&start=10 "Indeed:Analytics Jobs in Toronto") and scrape unsponsored job titles, companies, and postings
- `later number` unique jobs were scraped from the search terms: `analytics`,`etc`....
- Jobs were passed from Python to R using [Feather](https://blog.rstudio.com/2016/03/29/feather/ "Feather: A Fast On-Disk Format for Data Frames for R and Python, powered by Apache Arrow")

```{r Import Data,echo=TRUE}
#list our data files
searches <- c("BI","analytics")
files <- list("results_Business Intelligence_Toronto+ON.feather",
              "results_analytics_Toronto+ON.feather")

#read and collapse to data frame
datalist <- lapply(files,function(x){read_feather(x)})
data <- bind_rows(datalist,.id="search")

#fix quotations in column names
names(data)
names(data) <- c("search","text","titles","urls")

#check if we have redundant jobs
sum(duplicated(data[,2:4]))

#reduce to distinct jobs and clean up search column
data <- data[!duplicated(data[,2:4]),]
data$search <- plyr::mapvalues(data$search,from=c("1","2"),to=searches)
head(data)


test <- read_feather("data science.feather")
test[1]

```


```{r Process unigrams Data}
#word analysis
data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     group_by(tokens) %>%
     filter(!tokens %in% stopwords("en")) %>%
     summarize(count=n()) %>%
     arrange(desc(count))

```

seems like some 'low value' words mixed in here. it'll take some work to get somewhere we'd like to be with this.


```{r Process bigrams}

data %>%
     unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
     group_by(tokens) %>%
     filter(!tokens %in% stopwords("en")) %>%
     summarize(count=n()) %>%
     arrange(desc(count))
```

It looks like our data is quite dirty. It'll be extremely taxing 


```{r Cleaner bigrams}
bigrams_sep <- data %>% 
     unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
     separate(col=tokens,into=c("word1","word2"),sep=" ") %>%
     filter(!word1 %in% stopwords("en"),
            !word2 %in% stopwords("en")) %>%
     unite(tokens,word1,word2,sep=" ") %>%
     count(tokens,sort=TRUE)

bigrams_sep
class(bigrams_sep)

```

```{r Word Frequencies}

```

```{r Convert to TDM}

```

```{r Word Cloud}

```

```{r Network Analysis}

```






