---
title: "The Toronto Analytics Job Landscape"
author: "Summerhill"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  md_document:
    variant: markdown_github
  html_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=10, fig.path='Figs/',
                      warning=FALSE,message=FALSE,error=FALSE,include=TRUE,echo=TRUE)

set.seed(42)

# a note on options: 
# include=TRUE means that output of the code is shown in the document (intended for graphics).
# echo=TRUE means the code is shown in the document (intended for code that might be interesting for a reader).
# message and warning are for the text for loading libraries or if a function fails
# echo=TRUE means the code is not shown in the final document
# http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html
# http://rmarkdown.rstudio.com/developer_parameterized_reports.html
# https://guides.github.com/features/mastering-markdown/

```
- This document was rendered last on `r Sys.Date()`

##*THIS PROJECT IS STILL UNDER CONSTRUCTION*
The intention will be to mask the code as the project approaches completion.

##Executive Summary
- This project was produced for the Text Analytics Workshop for the Winter 2018 Masters of Management Analytics Cohort at Queen's University
- The goal from the outset was to use text analytics techniques developed in class to examine jobs companies have posted on Indeed in Toronto
and employ techniques discussed in class including document clustering, topic modelling, and visualization.

##Project Rationale
- A open sourced project working with real world data was desired
- Other projects can be found scraping DS/Analytics jobs from Indeed. Typically word frequencies for keywords like Python or Hadoop are calculated
- Moving beyond that, we were interested in topic modelling and how the choice of words signals relationships between roles
- Job postings fit the 'bag of words' or ngram approach taught in class. Not many employers say **"We don't want someone who knows Python"**

```{r Import Libraries}
library(feather)
library(tidyverse)
library(tidytext)
library(tm)
library(wordcloud)
library(widyr)
library(ggraph)
library(igraph)
library(knitr)
library(ggridges)
```

##Gathering Data
- Beautiful Soup & Selenium were used in Python to access [Indeed](https://www.indeed.ca/jobs?q=analytics&l=Toronto&start=10 "Indeed:Analytics Jobs in Toronto") and scrape unsponsored job titles, companies, and postings
- `later number` unique jobs were scraped from the search terms: `analytics`,`etc`....
- Jobs were passed from Python to R using [Feather](https://blog.rstudio.com/2016/03/29/feather/ "Feather: A Fast On-Disk Format for Data Frames for R and Python, powered by Apache Arrow")

```{r Import Data,echo=TRUE}
rm(list=ls())
#list our data files
searches <- c("analytics",
                 "data analyst",
                 "data scientist",
                 "analytics strategy",
                 "data insights",
                 "marketing analytics",
                 "analytics reporting",
                 "machine learning",
                 "business intelligence")

files <- paste("data/feather/",searches,".feather",sep="")

#read and collapse to data frame
datalist <- lapply(as.list(files),function(x){read_feather(x)})
test <- datalist[[1]]
data <- bind_rows(datalist,.id="search")
rm(datalist)

#fix quotations in column names
names(data) <- c("search","company","text","titles","urls")
data <- data %>% select(company,titles,text,search,urls)

#check if we have redundant jobs
sum(duplicated(data[,2:4]))

#examine the uniqueness of our data
NumJobs <- n_distinct(data$urls)

#reduce to distinct jobs and clean up search column
data <- data[!duplicated(data$urls),]
data$search <- plyr::mapvalues(data$search,
                               from=unique(data$search),
                               to=searches)
```
- Our data returned `NumJobs` unique jobs within our search.
- It's clear a considerable amount of cleaning is in order 

```{r preprocessing}
RemovePattern <- function(vector,pattern){gsub(pattern=pattern,replacement=" ",vector)}
data <- dmap(data,RemovePattern,"\n")
data <- dmap(data,RemovePattern,"\\(")   
data <- dmap(data,RemovePattern,"\\)")
data <- dmap(data,RemovePattern,"\\{[a-zA-Z0-9]\\}")
```

```{r Jobs Found}
#investigate redundant jobs. Should return 200/each if they are all unique.

rollup <- data %>%
     group_by(search) %>%
     summarize(NumberUniquePostings=n())

str(rollup)

#sort by search order
kable(left_join(data.frame(search=searches),rollup,by="search"))
```

- We expect 200 jobs for each result, and removing the duplicate jobs in the order they were searched.
- Interestingly, searching 200 jobs in analytics returns only 113 unique jobs, some redundancy exists.
- As we search overlapping terms, data sciencist, data insights, fewer and fewer unique jobs are returned
- Interestingly, each additional search term returns a surprising amount of new jobs. 75 jobs are shown for machine learning that were not found for data scientist, a fairly similar field.
- Business Intelligence seems to be fairly lateral to other search terms, returning many unique jobs

```{r}
kable(head(data %>% group_by(search,titles) %>% count(sort=TRUE),10))
```
```{r}
kable(head(data %>% count(company,sort=TRUE),10))
```

```{r Remove empty jobs}
#how long are our jobs
data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(urls) %>%
     count(search,urls,sort=TRUE) %>%
     filter(n<3000) %>%
     ggplot(aes(x=n,y=search))+geom_density_ridges()
```

- We see that there are alot of 0 information jobs, postings with only a few words, let's look at what some of those are.

```{r remove outliers}
MinWords <- 300

empty_urls <- data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(urls) %>%
     count(urls,sort=TRUE) %>%
     filter(n < MinWords)

# data %>% 
#      filter(urls %in% empty_urls$urls) %>%
#      head()

data <- data %>% 
     filter(!urls %in% empty_urls$urls)
     
```

- We see here a variety of failed scrapings, or 0 info postings


```{r long postings}
long_postings <- data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(urls) %>%
     count(urls,sort=TRUE) %>%
     filter(n >3000)

```

```{r}
unigrams <- data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(unigrams) %>%
     count(unigrams,sort=TRUE)

wordcloud(unigrams$unigrams,unigrams$n,max.words=50)
```


- Looking at a simple word frequency, we see out of the box our data is very messy
- The boiler plate at the end of each job posting, encouraging people to apply, discussing company acolades and culture distort our analysis. Let's spend some time cleaning up *common stopwords*, *job specific words* and *html*

```{r Process unigrams Data}
#what words to avoid
stop <- read.csv("stopwords.csv",stringsAsFactors = FALSE)
stop <- rbind(stop,data.frame(words=stopwords("en")))


#process n-grams
unigrams <- data %>%
     unnest_tokens(token="words",output="unigrams",input=text) %>%
     group_by(unigrams) %>%
     filter(!unigrams %in% stop$words) %>%
     count(unigrams,sort=TRUE)

#visualize
wordcloud(unigrams$unigrams,unigrams$n,max.words=50)
```

- We are starting to look better. Let's take a look at our bigrams.
```{r Process bigrams}
#look a bi-grams
bigrams_totals <- data %>%
     unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
     separate(col=tokens,into=c("word1","word2"),sep=" ") %>%
     filter(!word1 %in% stop$words, !word2 %in% stop$words) %>%
     unite(tokens,word1,word2,sep=" ") %>%
     count(tokens,sort=TRUE)

#kable(head(bigrams_totals,20))

wordcloud(bigrams_totals$tokens,bigrams_totals$n,max.words=10)
```



```{r Unigram frequency}

#determine how frequent each word occurs across job postings. Don't skip stop words yet.
unigrams_freq <- data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     select(urls,tokens) %>%
     distinct() %>%
     filter(!tokens %in% stop$words) %>%
     group_by(tokens) %>%
     count(tokens,sort=TRUE) %>%
     ungroup() %>%
     mutate(frequency=n/NumJobs)

kable(head(unigrams_freq,20))
```


```{r}

#determine how frequent each bigram is across job postings. Don't skip stop words yet.
bigrams_freq <- data %>%
     unnest_tokens(token="ngrams",n=2,output="tokens",input=text) %>%
     select(urls,tokens) %>%
     distinct() %>%
     separate(col=tokens,into=c("word1","word2"),sep=" ") %>%
     filter(!word1 %in% stop$words, !word2 %in% stop$words) %>%
     unite(tokens,word1,word2,sep=" ") %>%
     group_by(tokens) %>%
     count(tokens,sort=TRUE) %>%
     ungroup() %>%
     mutate(frequency=n/NumJobs)

kable(head(bigrams_freq,20))

```



```{r pairwise count}

skills <- c("sas","python","sql","r","spss",
            "hadoop","spark","java","scala","aws","c++",
            "excel","powerpoint","mapreduce",
          "qlik","tableau","microstrategy","azure")

data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     filter(tokens %in% skills) %>%
     pairwise_count(tokens,urls,sort=TRUE) %>%
     head()

data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     filter(tokens %in% skills) %>%
     ggplot(aes(x=tokens)) +
     geom_bar()
```





```{r pairwise correlation}
#pairwise correlation
data %>%
     unnest_tokens(token="words",output="tokens",input=text) %>%
     filter(tokens %in% skills) %>%
     pairwise_cor(tokens,urls,sort=TRUE) %>%
     filter(correlation > .2) %>%
     graph_from_data_frame() %>%
     ggraph(layout = "fr") +
     geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
     geom_node_point(color = "lightblue", size = 5) +
     geom_node_text(aes(label = name), repel = TRUE) +
     theme_void()

```

```{r clustering}

#creating tokens - once we come up with list of stop words, should use anti_join here to remove them
# 
# data_token <- data %>%
#      unnest_tokens(word,text)
# 
# #creating unique document identifiers for each job posting
# data_token$ID <- paste(data_token$company,data_token$titles,sep="_")
# 
# #removing useless columns -
# data_token <- data_token[,-(1:4)]
# head(data_token)
# 
# #frequency of words by document
# 
# data_count <- data_token %>%
#      count(ID,word,sort=TRUE)
# 
# #creating table of total count of words in each document
# data_count2 <- data_count %>%
#      group_by(ID) %>%
#      summarise(total = sum(n))
# 
# #joining the total count information with primary dataset
# data_count <- right_join(data_count,data_count2)
# 
# #creating the correlation column so we can filter out words <1% or >80% of docs
# data_count <- data_count %>%
#      mutate(correlation = n/total) %>%
#      arrange(desc(correlation))
# 
# #Note for Gage: I'm noticing something odd for some of these postings. Look at the head of the data frame. The Brain Hunter Systems job posting only has 2 words but when I go to the actual posting itself, it clearly has way more than that
# 
# head(data_count)
# 
# 
# #Correlation filters
# minCor = 0.01
# maxCor = 0.80
# 
# #filter out words <1% or >80% of docs
# data_count <- data_count %>%
#      filter(correlation > minCor & correlation < maxCor)
# 
# #casting tidy text data into DTM
# data_dtm <- data_count %>%
#      cast_dtm(ID,word,n)
# 
# #creating TDM
# data_tdm <- t(data_dtm)
# 
# #reducing sparsity of TDM before clustering
# dim(data_tdm)
# data_clean <- removeSparseTerms(data_tdm,sparse = 0.80)
# data_m <- as.matrix(data_clean)
# data_df <- as.data.frame(data_m)
# 
# #looking at most popular words, min freq of 20
# findFreqTerms(data_tdm,lowfreq=20)
# 
# #accumulator of clustering results
# clust_results <- data.frame()
# 
# #run kmeans for all clusters up to 15
# set.seed(50)
# for(i in 1:10) {
#      k_clust <- kmeans(data_m, centers=i, iter.max =100)
#      #Combine cluster number and cluster together, write to df
#      clust_results <- rbind(clust_results,cbind(i,k_clust$tot.withinss))
# }
# names(clust_results) <- c("cluster","results")
# 
# #scree elbow plot
# ggplot(clust_results,aes(x=cluster,y=results,group=1))+
# geom_point()+geom_line()+scale_x_continuous(name='# of clusters',limits = c(1, 10))
# 
# #if you want specific points look at scaling the axes then use seq(1:10)
# 
# ## above plot shows that at about 3 clusters, any additional clusters would lose significance
# 
# #need to create principal components, regular kmeans will kick out an error because it uses princomp function
# pc <- prcomp(data_m)
# plot(pc,xlab="Principal Components")
# summary(pc)
# comp <- data.frame(pc$x[,1:12])
# plot(comp)
# 
# #Let's look at 3 clusters
# kmeans3 <- kmeans(comp,centers=3)
# 
# #Plot clusters
# library(cluster)
# clusplot(comp, kmeans3$cluster, color=TRUE, shade=TRUE,
#          labels=2, lines=0, main="K-means cluster plot")
# 
# ###hierarchical clustering
# 
# #calculate distance
# data_dist <- dist(data_df,method="euclidean")
# 
# #creating and plotting hierarchical cluster
# hc <- hclust(data_dist)
# plot(hc)
# rect.hclust(hc,k=8,border="red")
```



```{r Clustering with dplyr}
# freqs_to_use <- data %>%
#      unnest_tokens(token="words",output="tokens",input=text) %>%
#      select(urls,tokens) %>%
#      distinct() %>%
#      filter(!tokens %in% stop$words) %>%
#      group_by(tokens) %>%
#      count(tokens,sort=TRUE) %>%
#      ungroup() %>%
#      mutate(frequency=n/NumJobs) %>%
#      filter(frequency>MinFreq & frequnecy<MaxFreq)
# 
# unis <- data %>%
#      unnest tokens %>%
#      select(-c(list of cols to drop)) %>%
#      filter(tokens %in% freqs_to_use$tokens) %>% #this takes care of stop words too
#      count() %>%
#      spread(tokens, but not the keys you use) # this should result in the features being words 
# # and the rows being different jobs. This is a dtm.
# 
# bigrams <-# same process. Look at the separate/unite functions for getting rid of stop words
# 
# datatocluster <- bind_cols(bigrams,unis)     
# #the result of this is a data frame where your rows are each job, maybe you need a key to identify
# # and each column is a unigram or bigram, and the data is how many times it was mentioned in the job
# 
# #do the same clustering
# 
# #when you now do this route, all your stop words should be gone. 
     
```







